---
title: "Travel Map"
author: "Ari Decter-Frain"
date: "2019-05-11"
description: "I wrote an R script to generate a map of my trips around the world. I used Google's Maps API to get geographic coordinates for the different cities and ggplot to create the map"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      fig.height = 5,
                      fig.width = 7.5,
                      dpi = 600,
                      message = FALSE)
```


I want to see as much of the world as possible while I can. The map I'm making here will serve as a reminder of the memories I made around the world, and as motivation to go out and see the rest. 

I've made the map in Rmarkdown so anyone reading who has a basic familiarity with R and Rstudio should be able to easily make their own.

To make the map I used R, ggplot, and Google's geolocation API. I relied heavily on [The Lucid Manager's](https://lucidmanager.org/create-air-travel-route-maps/) post in which he creates his own version. 

<hr>

First loading all the necessary libraries
```{r}
library(tidyverse)
library(ggmap)
library(ggrepel)
library(maps)
library(here)
```


Here I'm setting internal filepaths. To reproduce the code below, replace the ``here()`` functions with the urls in the comments.
```{r}
path_flights <- here("static", "map", "trips.csv")
#https://raw.githubusercontent.com/aridf/travel-map/master/data/trips.csv
path_geos <- here("static", "map", "geos.csv")
#https://raw.githubusercontent.com/aridf/travel-map/master/data/geos.csv
```


Loading the data. I had to add country info to some of the cities to ensure the google API searches for the correct city. I remove this later so it does not show up in my visualization.
```{r}
flights <- read_csv(path_flights)
head(flights)
```


Here I'm deleting duplicates and return flights. This helps reduce the clutter in the visual.
```{r message = FALSE}
d <- vector()
for (i in 1:nrow(flights)) {
    d <- which(paste(flights$From, flights$To) 
               %in% paste(flights$To[i], flights$From[i]))
    flights$From[d] <- "R"
}
flights <- flights %>%
  filter(From != "R") %>%
  select(From, To)
```


Next, I grab coordinate data from the google API. I use two layers of conditions to limit the number of calls I'm making to the API. First, I check if I already have a list of locations and coordinates saved. If not, I get a list of all the locations in my flight list and call that API for each. If I already have some coordinates saved, I just check to see if there are any new locations. If there are, I call the API for the new entries only. If not, then I move ahead with my previous dataset. 
```{r message = FALSE}
locations <- unique(c(flights$From, flights$To))
if(file.exists(path_geos)) {
  geos <- read_csv(path_geos)
  new <- locations[!(locations %in% geos$airport)]
  if(length(new) > 0) {
    geos <- new %>%
      geocode() %>%
      mutate(airport = new) %>%
      bind_rows(geos) %>%
      select(airport, lon, lat)
    write_csv(geos, path_geos)
  }
} else {
  new <- unique(c(flights$From, flights$To))
  geos <- new %>%
    geocode() %>%
    mutate(airport = new) %>%
    select(airport, lon, lat)
  write_csv(geos, path_geos)
}
head(geos)
```


Merging trip data and coordinates.
```{r message = FALSE}
flights <- merge(flights, geos, by.x = "To", by.y = "airport")
flights <- merge(flights, geos, by.x = "From", by.y = "airport")
head(flights)
```


Removing country info from the city names here.
```{r}
geos$airport <- geos$airport %>%
  word(1, sep = ",")
```



Creating my world map object. I'm setting the size limits of my map dynamically so it will expand as I travel to more continents. 
```{r message = FALSE}

#set size limits of the worldmap. Prevent unvisited regions from displaying
xmin <- min(geos$lon) - 10
xmax <- max(geos$lon) + 10
ymin <- min(geos$lat) - 10
ymax <- max(geos$lat) + 10

#get map
worldmap <- borders("world", xlim = c(xmin, xmax), ylim = c(ymin, ymax), 
                    colour = "#eee8d5", fill = "#eee8d5") 
```


This is the actual plot. The layers are pretty straightforward:

* `` geom_curve `` creates curved lines connecting "From" and "To".
* `` geom_point `` adds dots at each location
* `` geom_text_repel `` add labels while giving them room to breathe 
```{r message = FALSE}
ggplot() + worldmap + 
    geom_curve(data = flights, aes(x = lon.x, y = lat.x, xend = lon.y, 
                                   yend = lat.y), col = "#2aa198", size = .4) + 
    geom_point(data = geos, aes(x = lon, y = lat), col = "#268bd2") + 
    geom_text_repel(data = geos, color = "#586e75", aes(x = lon, y = lat, label = airport), 
               col = "black", size = 2, segment.color = NA, segment.size = 1) + 
    theme_void() +
    theme(panel.background = element_rect(fill = "#fdf6e3"))
```

